A Tale of Two Transects

### 2020-02-03 writing/transect_paper.Rmd
Clean up figures after addition of Illumina data.

An additional sample was collected at every sampling location in 2016 using an older batch of lysis buffer for preservation.
Sequencing results for these samples were found not to be comparable with results from the newer batch, so these samples were not included in spatial analyses.
However, reads from these sequences were included in the full bioinformatics workflow, including ASV calling, OTU clustering, and phylogenetic trees.

@@

After field lysis, DNA was extracted using the Xpedition^TM^ Soil/Fecal Prep kit (see above).
DNA was quantified using fluorometrically using Quant-iT^TM^ PicoGreen^TM^ dsDNA (Thermo Fisher Scientific, Waltham, MA, USA) fluorescent indicator dye on a Infinite F200 plate spectrofluorometer (Tecan Trading AG, Männedorf, Switzerland) according to the manufacturer's protocol.

@@

venn_ASV %>%
  filter(!is.na(ASVs)) %>%
  set_names(gsub("_[[:alpha:]]+[-._]\\d+$", "", names(.))) %>%
  set_names(gsub("_", " ", names(.))) %>%
  kable(
    booktabs = TRUE,
    caption = "(ref:venn-table-asv)",
    linesep = c(rep("", 2), "\\addlinespace", rep("", 2), "\\addlinespace"),
    align = ("rcrrcrrcrr"),
    row.names = FALSE
  ) %>%
  kableExtra::add_header_above(
    header = c(
      " " = 1,
      names(venn_ASV)[-(1)] %>% 
        gsub(pattern = "(ASV_)?(reads|frac)_", replacement = "") %>%
        rle() %$%
        set_names(lengths, plyr::mapvalues(values, datasets$seq_run, datasets$amplicon, FALSE))
      )
  ) %>%
  kableExtra::add_header_above(
    header = c(
      " " = 1,
       names(venn_ASV)[-(1)] %>% 
        gsub(pattern = "(ASV_)?(reads|frac)_", replacement = "") %>%
        plyr::mapvalues(datasets$seq_run, paste(datasets$tech, datasets$machine), FALSE) %>%
        rle() %$%
        set_names(lengths, values)
      )
  )
```

@@

venn_OTU %>%
  set_names(gsub("_[[:alpha:]]+[-._]\\d+$", "", names(.))) %>%
  set_names(gsub("_", " ", names(.))) %>%
  kable(
    booktabs = TRUE,
    caption = "(ref:venn-table-otu)",
    linesep = c(rep("", 2), "\\addlinespace", rep("", 2), "\\addlinespace"),
    align = ("rcrrcrrcrr"),
    row.names = FALSE
  ) %>%
  kableExtra::add_header_above(
    header = c(
      " " = 1,
      names(venn_OTU)[-1] %>% 
        gsub(pattern = "(OTU_)?(reads|frac)_", replacement = "") %>%
        rle() %$%
        set_names(lengths, plyr::mapvalues(values, datasets$seq_run, datasets$amplicon, FALSE))
      )
  ) %>%
  kableExtra::add_header_above(
    header = c(
     " " = 1,
      names(venn_OTU)[-1] %>% 
        gsub(pattern = "(OTU_)?(reads|frac)_", replacement = "") %>%
        plyr::mapvalues(datasets$seq_run, paste(datasets$tech, datasets$machine), FALSE) %>%
        rle() %$%
        set_names(lengths, values)
      )

@@

```{r venn, fig.cap = "(ref:venn)", out.height = "30%", fig.show="hold"}
vennplot_ASV <- venn_ASV %>%
  rownames_to_column("set") %>%
  mutate(set = map_chr(set,
                       ~ strsplit(., "")[[1]] %>%
                     as.integer() %>%
                     as.logical() %>%
                     magrittr::extract(LETTERS[seq_along(.)], .) %>%
                     paste(collapse = "&"))) %>%
  mutate(
    reads_frac = paste(frac_pb_500, frac_pb_483, frac_is_057, `frac_SH-2257`,
                       sep = "/") %>%
      gsub("/+", "/", .) %>%
      gsub("(^/|/$)", "", .),
    label = paste(ASVs, reads_frac, sep = "\n")
  )

@@

vennplot_OTU <- venn_OTU %>%
  rownames_to_column("set") %>%
  mutate(n = str_count(set, "1")) %>%
  arrange(n, desc(set)) %>%
  mutate(set = map_chr(set,
                       ~ strsplit(., "")[[1]] %>%
                     as.integer() %>%
                     as.logical() %>%
                     magrittr::extract(LETTERS[seq_along(.)], .) %>%
                     paste(collapse = "&"))) %>%
  mutate(
    reads_frac = paste(frac_pb_500, frac_pb_483, frac_is_057, `frac_SH-2257`,
                       sep = "/") %>%
      gsub("/+", "/", .) %>%
      gsub("(^/|/$)", "", .),
    label = paste(OTUs, reads_frac, sep = "\n")
  )

eulerr::venn(deframe(select(vennplot_OTU, set, OTUs))) %>%
  plot(
    legend = FALSE,
    labels = FALSE,
    quantities = vennplot_OTU$label,
    main = "OTU"
  )

@@

read_comparison <- function(multi_table, comparisons, type) {
  multi_table <- filter(multi_table, type == !!type)
  comparisons <- filter(comparisons, type == !!type)
  p <- multi_table %>%
    filter(x > 0, y > 0) %>%
    ggplot(aes(y = y, x = x)) +
    geom_point(alpha = 0.2, shape = 1) +
    scale_x_log10() +
    scale_y_log10() +
    coord_equal() +
    xlab(NULL) +
    ylab(NULL) +
    geom_rug(aes(y = y), sides = "l", alpha = 0.2,
             data = multi_table %>% filter(x == 0, y > 0)) +
    geom_rug(aes(x = x), sides = "b", alpha = 0.2,
             data = multi_table %>% filter(y == 0, x > 0)) +
    facet_grid(y_var ~ x_var, switch = "both", space = "free") +
    stat_smooth(method = "loess", formula = y ~ x) +
    geom_abline(aes(intercept = `(Intercept)`, slope = 1),
                linetype = 2L, alpha = 0.5, data = comparisons) +
    geom_text(
      aes(label = label),
      data = comparisons,
      x = 0.2,
      y = 5.5,
      color = "black",
      hjust = 0,
      parse = TRUE,
      inherit.aes = FALSE
    ) +
    theme(strip.placement = "outside", strip.background = element_blank())
  g <- ggplotGrob(p)
  g$grobs[g$layout$name %in% c("panel-1-2", "panel-1-3", "panel-2-3")] <- NULL
  g$layout <- g$layout[!(g$layout$name %in% c("panel-1-2", "panel-1-3", "panel-2-3")),]
  g
}
ggpubr::ggarrange(
  read_comparison(multi_table, comparisons, type = "ASV"),
  read_comparison(multi_table, comparisons, type = "OTU"),
  ncol = 1,
  labels = "auto",
  label.x = 0.9
)

read_comparison <- function(big_table, comparisons, x, y, xlab, ylab) {
  x_var <- enquo(x)
  y_var <- enquo(y)
  big_table %>%
  filter(!!x_var > 0, !!y_var > 0) %>%
  ggplot(aes(x = !!x_var, y = !!y_var)) +
  geom_point(alpha = 0.2, shape = 1) +
  scale_x_log10() +
  scale_y_log10(limits = c(1, 1e4)) +
  coord_equal() +
  xlab(xlab) +
  ylab(ylab) +
  geom_rug(aes(y = !!y_var, x = 1), sides = "l", alpha = 0.2,
           data = big_table %>% filter(!!x_var == 0, !!y_var > 0)) +
  geom_rug(aes(y = !!x_var), sides = "b", alpha = 0.2,
           data = big_table %>% filter(!!y_var == 0, !!x_var > 0)) +
  stat_smooth(method = "loess", formula = y ~ x) +
  geom_abline(aes(intercept = mean(log10((!!y_var)[(!!y_var) >= 1])) -
                    mean(log10((!!x_var)[(!!x_var) >= 1])), slope = 1),
              linetype = 2L, alpha = 0.5) +
  geom_text(
    aes(label = label),
    data = filter(comparisons, .data$x == quo_name(x_var), .data$y == quo_name(y_var)),
    x = 1,
    y = 3.5,
    color = "black",
    hjust = 0,
    parse = TRUE,
    inherit.aes = FALSE
  ) +
  facet_wrap(~type, strip.position = "right", scales = "free")
}
```

@@

```{r tax-chart, fig.cap = "(ref:tax-chart)"}
readd(taxon_reads, cache = cache) %>%
  pivot_longer(names_to = "rank", cols = c("kingdom", "phylum", "class", "order", "family", "genus"), values_to = "taxon") %>%
  group_by(amplicon, tech, Algorithm, reference, rank, ID = !is.na(taxon)) %>%
  summarize(reads = sum(reads), ASVs = n()) %>%
  group_by(amplicon, tech, Algorithm, reference, rank) %>%
  mutate(ASVs = ASVs / sum(ASVs)) %>%
  ungroup() %>%
  filter(ID) %>%
  mutate_at("rank", factor, levels = c("kingdom", "phylum", "class", "order", "family", "genus")) %>%
  rename(Reads = reads) %>%
  mutate_at("reference", as.character) %>%
  mutate(
    reference = map(reference, ~ if (. == "All") c("Unite", "Warcup", "RDP") else .)
  ) %>%
  unnest(reference) %>%
  mutate_at("reference", factor, levels = c("Unite", "Warcup", "RDP")) %>%
  pivot_longer(cols = c("Reads", "ASVs"), names_to = "type", values_to = "frac") %>%
  ggplot(aes(x = rank, y = frac, ymax = frac, group = Algorithm, fill = Algorithm)) +
  ggnomics::facet_nested(
    reference ~ type + tech + amplicon,
    resect = unit(1, "mm"),
    nest_line = TRUE,
    bleed = FALSE
  ) +
  # geom_ribbon(ymin = 0, alpha = 0.5) +
  # geom_line(aes(color = Algorithm)) +
  geom_col(position = "dodge") +
  ylab("Fraction of reads assigned") +
  xlab(NULL) +
  theme(axis.text.x = element_text(vjust = 0.5, angle = 90),
        legend.direction = "horizontal",
        legend.position = "bottom",
        strip.background = element_blank())
```

@@

```{r taxa, fig.cap = "(ref:taxa)", out.width="100%", fig.width = 8}
set.seed(2)
readd(taxdata, cache = cache) %>%
  metacoder::heat_tree(
    layout = "davidson-harel",
    # initial_layout = "reingold-tilford",
    node_size = rowMeans(.$data$tax_read[,-1]),
    node_color = rowMeans(.$data$tax_read[,-1]),
    node_label = replace_na(taxon_names, "unknown"),
    node_size_range = c(0.002, .035),
    node_size_axis_label = "Read abundance",
    node_color_axis_label = "Read abundance",
    node_label_size_range = c(0.01, 0.03),
    edge_size = rowMeans(.$data$tax_asv[,-1]),
    edge_size_trans = "linear",
    edge_size_axis_label = "ASV count",
    edge_size_range = c(0.001, 0.03),
    edge_color = rowMeans(.$data$tax_asv[,-1]),
    edge_color_axis_label = "ASV count",
    # aspect_ratio = 3/2,
    make_node_legend = FALSE,
    make_edge_legend = FALSE,
    output_file = here("temp/heattree.pdf")
  )
```

@@

set.seed(2)
taxdata2 %>%
  metacoder::heat_tree(
    layout = "davidson-harel",
    initial_layout = "reingold-tilford",

@@

    node_label = ifelse(
      pmax(abs(.$data$diff_read$read_ratio),
           abs(.$data$diff_asv$asv_ratio)) > 0.5,
      taxon_names,
      ""
    ),
    node_label_size_range = c(0.010, 0.03),
    # title = unique(paste(.$data$diff_read$treatment_1, "vs.", .$data$diff_read$treatment_2)),
    aspect_ratio = 3/2
  )
```

@@

(ref:correlog-cap) Mantel correlograms showing the correlation between Bray-Curtis (top two rows) or weighted Unifrac (bottom row) dissimilarities and spatial distance, for short (top row) and long (bottom two rows) amplicons, and for the entire soil fungal community (left) or only ECM taxa (right).  Unifrac distance was not calculated for short amplicons because it requires that the sequences are placed on a phylogenetic tree.

@@

light_plan %>%
  filter(startsWith(target, "correlog")) %>%
  select(target) %>%
  separate(
    target,
    into = c("step", "guild", "metric", "tech", "amplicon", "tdist"),
    sep = "_",
    remove = FALSE
  ) %>%
  mutate(
    tdist = paste(tdist, "year"),
    guild = plyr::mapvalues(guild, c("fungi", "ecm"), c("all Fungi", "ECM")),
    metric = plyr::mapvalues(metric, c("bray", "wunifrac"), c("Bray-Curtis", "Weighted UNIFRAC")),
    amplicon = factor(amplicon, levels = c("short", "long"), labels = c("short amplicon", "long amplicon"))
  ) %>%
  mutate_at("target", purrr::map, readd, character_only = TRUE, cache = cache) %>%
  mutate_at("target", purrr::map, "mantel.res") %>%
  mutate_at("target", purrr::map, as.data.frame) %>%
  unnest(target) %>%
  filter(complete.cases(.)) %>%
  mutate(Significant = `Pr(corrected)` < 0.05) %>%
  ggplot(aes(class.index, Mantel.cor, group = tdist, color = tdist, shape = Significant)) +
  geom_point(size = 3) +
  geom_line(alpha = 0.3, size = 1.5) +
  geom_hline(yintercept = 0, linetype = 2, alpha = 0.5) +
  # geom_line(aes(y = exp(predict(bc_fit)))) +
  scale_y_continuous(limits = c(-0.05, NA), name = "Mantel correlation") +
  scale_x_continuous(breaks = c(0, 3, 6, 9, 12)) +
  scale_color_discrete(name = "Time lag") +
  xlab("Distance (m)") +
  ggnomics::facet_nested(metric + tech + amplicon ~ guild, scales = "fixed") +
  scale_shape_manual(values = c("TRUE" = 16, "FALSE" = 1))
```

@@

```{r spatio-temporal, fig.cap = "(ref:spatio)"}
variog_data <- light_plan %>%
  filter(step == "variofitST") %>%
  select(variofitST, variogST, guild, metric, tech, amplicon) %>%
  mutate_at(c("guild", "metric", "tech", "amplicon"), str_replace_all, "\"", "") %>%
  mutate(
    guild = plyr::mapvalues(guild, c("fungi", "ecm"), c("all Fungi", "ECM")),
    metric = plyr::mapvalues(metric, c("bray", "wunifrac"), c("Bray-Curtis", "Weighted UNIFRAC")),
    amplicon = factor(
      amplicon,
      levels = c("short", "long"),
      labels = c("short amplicon", "long amplicon")
    )
  ) %>%
  mutate_at(
    c("variogST", "variofitST"),
    lapply,
    readd,
    character_only = TRUE,
    cache = cache)

@@

variog_empir %>%
  ggplot(aes(dist, gamma, color = timelag, size = n)) +
  # geom_hline(data = variog_sills, aes(yintercept = sills), linetype = "dashed", color = "gray50") +
  # geom_vline(data = variog_data, aes(xintercept = range), linetype = "dashed", color = "gray50") +
  # geom_jitter(shape = 46, alpha = 0.5) +
  # geom_line(size = 1.5, alpha = 0.5) +
  geom_point(alpha = 0.5) +
  # geom_smooth(data = variog_points, aes(fill = timelag), color = FALSE, alpha = 0.5) +
  geom_line(data = filter(variog_fit, timelag == 1), size = 0.5, color = "darkcyan", alpha = 0.8) +
  geom_line(data = filter(variog_fit, timelag == 0), size = 0.5, color = "red1", alpha = 0.8) +
  scale_x_continuous(expand = expand_scale(0, 0), limits = c(0, 25), name = "Distance (m)") +
  scale_color_discrete(name = "Time lag") +
  scale_y_continuous(name = "Community dissimilarity") +
  ggnomics::facet_nested(metric + tech + amplicon ~ guild, scales = "free")

@@

(ref:full-length) Comparison of the length of sequences from different length amplicons and sequencing technologies.  *a*) Length distribution.  *b*) Empirical cumulative distribution function.

```{r full-length, fig.cap = "(ref:full-length)", fig.height = 6}
full_dens <- 
  readd(reads_table, cache = cache) %>%
  dplyr::left_join(dplyr::select(datasets, seq_run, tech)) %>%
  dplyr::filter(region %in% c("long", "short")) %>%
  dplyr::group_by(seq_run) %>%
  dplyr::mutate(
    amplicon = ifelse(seq_run == "pb_500", "Long", "Short") %>%
      factor(levels = c("Short", "Long")),
    reads = reads / sum(reads)
  ) %>%
  ggplot(aes(length, weight = reads, color = seq_run, group = seq_run)) +
  stat_density(bw = 0.5, geom = "line", position = "identity") +
  scale_color_brewer(type = "qual", guide = "none", palette = "Set2") +
  scale_y_continuous(name = "Density") +
  scale_x_continuous(name = NULL) +
  facet_wrap(~tech + amplicon, ncol = 1, strip.position = "right", scales = "free_y")

full_ecdf <- readd(reads_table, cache = cache) %>%
  dplyr::filter(region  %in% c("long", "short")) %>%
  dplyr::group_by(seq_run) %>%
  dplyr::mutate(reads = reads / sum(reads)) %>%
  ggplot(aes(length, weight = reads, color = seq_run, group = seq_run)) +
  stat_ecdf(geom = "line") +
  scale_color_brewer(type = "qual",
                     labels = c("Short - Ion", "Short - PacBio", "Long - PacBio", "Short - Illumina"),
                     name = NULL, palette = "Set2") +
  scale_y_continuous(name = "Cumulative read fraction") +
  scale_x_continuous(name = "Length (bp)") +
  theme(legend.position = c(1, 0.02), legend.justification = c(1, 0), legend.background = element_blank())

ggpubr::ggarrange(full_dens, full_ecdf, ncol = 1, labels = "auto")
```

>>>>

An additional sample was collected at every sampling location (1-m spacing) in 2016 using LifeGuard^TM^ Soil Preservation Solution (MO BIO, Carlsbad, CA; USA) for preservation, without field lysis.
Sequencing results for these samples were found not to be comparable with results preserved using the Xpedition^TM^ lysis buffer (Supplementary figure \@ref(fig:)), so these samples were not included in spatial analyses.
However, reads from these sequences were included in the full bioinformatics workflow, including ASV calling, OTU clustering, and phylogenetic trees.

@@

After field lysis, DNA was extracted using the Xpedition^TM^ Soil/Fecal Prep kit (see above).
Samples preserved using 
DNA was quantified using fluorometrically using Quant-iT^TM^ PicoGreen^TM^ dsDNA (Thermo Fisher Scientific, Waltham, MA, USA) fluorescent indicator dye on a Infinite F200 plate spectrofluorometer (Tecan Trading AG, Männedorf, Switzerland) according to the manufacturer's protocol.

@@

linesep <- venn_ASV %>%
  filter(!is.na(ASVs)) %>%
  select(starts_with("reads")) %>%
  mutate_all(nchar) %>%
  as.matrix %>%
  is_greater_than(0) %>%
  rowSums %>%
  rle() %$%
  map(lengths, ~ c(rep("", . - 1), "\\addlinespace")) %>%
  unlist() %>%
  head(length(.) - 2) %>%
  c("\\midrule")

venn_ASV %>%
  filter(!is.na(ASVs)) %>%
  set_names(gsub("_[[:alpha:]]+[-._]\\d+$", "", names(.))) %>%
  set_names(gsub("_", " ", names(.))) %>%
  kable(
    booktabs = TRUE,
    caption = "(ref:venn-table-asv)",
    linesep = linesep,
    align = ("rcrrcrrcrr"),
    row.names = FALSE
  ) %>%
  kableExtra::add_header_above(
    header = c(
      " " = 1,
      names(venn_ASV)[-(1)] %>% 
        gsub(pattern = "(ASV_)?(reads|frac)_", replacement = "") %>%
        rle() %$%
        set_names(lengths, plyr::mapvalues(values, datasets$seq_run, datasets$amplicon, FALSE))
      )
  ) %>%
  kableExtra::add_header_above(
    header = c(
      " " = 1,
       names(venn_ASV)[-(1)] %>% 
        gsub(pattern = "(ASV_)?(reads|frac)_", replacement = "") %>%
        plyr::mapvalues(datasets$seq_run, paste(datasets$tech, datasets$machine), FALSE) %>%
        rle() %$%
        set_names(lengths, values)
      )
  )
```

@@

linesep <- venn_OTU %>%
  filter(!is.na(OTUs)) %>%
  select(starts_with("reads")) %>%
  mutate_all(nchar) %>%
  as.matrix %>%
  is_greater_than(0) %>%
  rowSums %>%
  rle() %$%
  map(lengths, ~ c(rep("", . - 1), "\\addlinespace")) %>%
  unlist() %>%
  head(length(.) - 2) %>%
  c("\\midrule")

venn_OTU %>%
  set_names(gsub("_[[:alpha:]]+[-._]\\d+$", "", names(.))) %>%
  set_names(gsub("_", " ", names(.))) %>%
  kable(
    booktabs = TRUE,
    caption = "(ref:venn-table-otu)",
    linesep = linesep,
    align = ("rcrrcrrcrr"),
    row.names = FALSE
  ) %>%
  kableExtra::add_header_above(
    header = c(
      " " = 1,
      names(venn_OTU)[-1] %>% 
        gsub(pattern = "(OTU_)?(reads|frac)_", replacement = "") %>%
        rle() %$%
        set_names(lengths, plyr::mapvalues(values, datasets$seq_run, datasets$amplicon, FALSE))
      )
  ) %>%
  kableExtra::add_header_above(
    header = c(
     " " = 1,
      names(venn_OTU)[-1] %>% 
        gsub(pattern = "(OTU_)?(reads|frac)_", replacement = "") %>%
        plyr::mapvalues(datasets$seq_run, paste(datasets$tech, datasets$machine), FALSE) %>%
        rle() %$%
        set_names(lengths, values)
      )

@@

```{r venn, fig.cap = "(ref:venn)", out.height = "30%", fig.show="hold"}
vennplot_ASV <- venn_ASV %>%
  rownames_to_column("set") %>%
  filter(set != "Total") %>%
  mutate(set = map_chr(set,
                       ~ strsplit(., "")[[1]] %>%
                     as.integer() %>%
                     as.logical() %>%
                     magrittr::extract(LETTERS[seq_along(.)], .) %>%
                     paste(collapse = "&"))) %>%
  mutate(
    reads_frac = paste(frac_pb_500, frac_pb_483, frac_is_057, `frac_SH-2257`,
                       sep = "/") %>%
      gsub("/+", "/", .) %>%
      gsub("(^/|/$)", "", .),
    label = paste(ASVs, reads_frac, sep = "\n")
  )

@@

vennplot_OTU <- venn_OTU %>%
  rownames_to_column("set") %>%
  filter(set != "Total") %>%
  mutate(set = map_chr(set,
                       ~ strsplit(., "")[[1]] %>%
                     as.integer() %>%
                     as.logical() %>%
                     magrittr::extract(LETTERS[seq_along(.)], .) %>%
                     paste(collapse = "&"))) %>%
  mutate(
    reads_frac = paste(frac_pb_500, frac_pb_483, frac_is_057, `frac_SH-2257`,
                       sep = "/") %>%
      gsub("/+", "/", .) %>%
      gsub("(^/|/$)", "", .),
    label = paste(OTUs, reads_frac, sep = "\n")
  )

eulerr::venn(deframe(select(vennplot_OTU, set, OTUs)) %>% replace_na(0)) %>%
  plot(
    legend = FALSE,
    labels = FALSE,
    quantities = vennplot_OTU$label,
    main = "OTU"
  )

@@

read_comparison <- function(multi_table, comparisons, type) {
  multi_table <- filter(multi_table, type == !!type)
  comparisons <- filter(comparisons, type == !!type)
  p <- multi_table %>%
    filter(x > 0, y > 0) %>%
    ggplot(aes(y = y, x = x)) +
    geom_point(alpha = 0.2, shape = 1) +
    scale_x_log10(
      breaks = c(1, 1000, 1000000),
      labels = c("1", "1k", "1M"),
      minor_breaks = c(10, 100, 10000, 100000),
      limits = c(1, 1e6)
    ) +
    scale_y_log10(
      breaks = c(1, 1000, 1000000),
      labels = c("1", "1k", "1M"),
      minor_breaks = c(10, 100, 10000, 100000),
      limits = c(1, 1e6)
    ) +
    coord_equal() +
    xlab(NULL) +
    ylab(NULL) +
    geom_rug(aes(y = y), sides = "l", alpha = 0.2,
             data = multi_table %>% filter(x == 0, y > 0)) +
    geom_rug(aes(x = x), sides = "b", alpha = 0.2,
             data = multi_table %>% filter(y == 0, x > 0)) +
    ggnomics::facet_nested(y_amplicon + y_var ~ x_amplicon + x_var, switch = "both", space = "free", nest_line = TRUE) +
    stat_smooth(method = "loess", formula = y ~ x) +
    geom_abline(aes(intercept = `(Intercept)`, slope = 1),
                linetype = 2L, alpha = 0.5, data = comparisons) +
    geom_text(
      aes(label = label),
      data = comparisons,
      x = 6,
      y = 0.7,
      color = "black",
      hjust = 1,
      parse = TRUE,
      inherit.aes = FALSE
    ) +
    theme(strip.placement = "outside", strip.background = element_blank())
  g <- ggplotGrob(p)
  g$grobs[g$layout$name %in% c("panel-1-2", "panel-1-3", "panel-2-3")] <- NULL
  g$layout <- g$layout[!(g$layout$name %in% c("panel-1-2", "panel-1-3", "panel-2-3")),]
  g
}
ggpubr::ggarrange(
  read_comparison(multi_table, comparisons, type = "ASV"),
  read_comparison(multi_table, comparisons, type = "OTU"),
  ncol = 1,
  labels = "auto",
  label.x = 0.9
)
```

@@

```{r tax-chart, fig.cap = "(ref:tax-chart)"}
readd(taxon_reads, cache = cache) %>%
  pivot_longer(names_to = "rank", cols = c("kingdom", "phylum", "class", "order", "family", "genus"), values_to = "taxon") %>%
  group_by(amplicon, tech, Algorithm, reference, rank, ID = !is.na(taxon)) %>%
  summarize(reads = sum(reads), ASVs = n()) %>%
  group_by(amplicon, tech, Algorithm, reference, rank) %>%
  mutate(ASVs = ASVs / sum(ASVs)) %>%
  ungroup() %>%
  filter(ID) %>%
  mutate_at("rank", factor, levels = c("kingdom", "phylum", "class", "order", "family", "genus")) %>%
  rename(Reads = reads) %>%
  mutate_at("reference", as.character) %>%
  mutate(
    reference = map(reference, ~ if (. == "All") c("Unite", "Warcup", "RDP") else .)
  ) %>%
  unnest(reference) %>%
  mutate_at("reference", factor, levels = c("Unite", "Warcup", "RDP")) %>%
  pivot_longer(cols = c("Reads", "ASVs"), names_to = "type", values_to = "frac") %>%
  ggplot(aes(x = rank, y = frac, ymax = frac, group = Algorithm, fill = Algorithm)) +
  ggnomics::facet_nested(
    tech + amplicon ~ type + reference,
    resect = unit(1, "mm"),
    nest_line = TRUE,
    bleed = FALSE
  ) +
  # geom_ribbon(ymin = 0, alpha = 0.5) +
  # geom_line(aes(color = Algorithm)) +
  geom_col(position = "dodge") +
  ylab("Fraction of reads assigned") +
  xlab(NULL) +
  theme(axis.text.x = element_text(vjust = 0.5, angle = 90),
        legend.direction = "horizontal",
        legend.position = "bottom",
        strip.background = element_blank())
```

@@

```{r taxa, fig.cap = "(ref:taxa)", out.width="100%", fig.width = 8}
set.seed(2)
theme <- theme_update(panel.border = element_blank())
readd(taxdata, cache = cache) %>%
  metacoder::heat_tree(
    layout = "davidson-harel",
    # initial_layout = "reingold-tilford",
    node_size = rowMeans(.$data$tax_read[,-1]),
    node_color = rowMeans(.$data$tax_read[,-1]),
    node_label = replace_na(taxon_names, "unknown"),
    node_size_range = c(0.002, .035),
    node_size_axis_label = "Read abundance",
    node_color_axis_label = "Read abundance",
    node_label_size_range = c(0.01, 0.03),
    edge_size = rowMeans(.$data$tax_asv[,-1]),
    edge_size_trans = "linear",
    edge_size_axis_label = "ASV count",
    edge_size_range = c(0.001, 0.03),
    edge_color = rowMeans(.$data$tax_asv[,-1]),
    edge_color_axis_label = "ASV count",
    # aspect_ratio = 3/2,
    make_node_legend = FALSE,
    make_edge_legend = FALSE,
    output_file = here("temp/heattree.pdf")
  )
theme_set(theme)
```

@@

set.seed(2)
theme <- theme_update(panel.border = element_blank())
taxdata2 %>%
  metacoder::heat_tree(
    layout = "davidson-harel",
    initial_layout = "reingold-tilford",

@@

    node_label = ifelse(
      pmax(abs(.$data$diff_read$read_ratio),
           abs(.$data$diff_asv$asv_ratio)) > 0.5,
      taxon_names,
      ""
    ),
    node_label_size_range = c(0.010, 0.03),
    # title = unique(paste(.$data$diff_read$treatment_1, "vs.", .$data$diff_read$treatment_2)),
    aspect_ratio = 3/2
  )
theme_set(theme)
```

@@

(ref:correlog-cap) Mantel correlograms showing the correlation between Bray-Curtis (top three rows) or weighted Unifrac (bottom row) dissimilarities and spatial distance, for short (top two rows) and long (bottom two rows) amplicons, and for the entire soil fungal community (left) or only ECM taxa (right).  Unifrac distance was not calculated for short amplicons because it requires that the sequences are placed on a phylogenetic tree.

@@

light_plan %>%
  filter(startsWith(target, "correlog")) %>%
  select(target) %>%
  separate(
    target,
    into = c("step", "guild", "metric", "tech", "amplicon", "tdist"),
    sep = "_",
    remove = FALSE
  ) %>%
  mutate(
    tdist = paste(tdist, "year"),
    guild = plyr::mapvalues(guild, c("fungi", "ecm"), c("all Fungi", "ECM")),
    metric = plyr::mapvalues(metric, c("bray", "wunifrac"), c("Bray-Curtis", "W-UNIFRAC")),
    amplicon = factor(amplicon, levels = c("Short", "Long"))
  ) %>%
  mutate_at("target", purrr::map, readd, character_only = TRUE, cache = cache) %>%
  mutate_at("target", purrr::map, "mantel.res") %>%
  mutate_at("target", purrr::map, as.data.frame) %>%
  unnest(target) %>%
  filter(complete.cases(.)) %>%
  mutate(Significant = `Pr(corrected)` < 0.05) %>%
  ggplot(aes(class.index, Mantel.cor, group = tdist, color = tdist, shape = Significant)) +
  geom_point(size = 3) +
  geom_line(alpha = 0.3, size = 1.5) +
  geom_hline(yintercept = 0, linetype = 2, alpha = 0.5) +
  # geom_line(aes(y = exp(predict(bc_fit)))) +
  scale_y_continuous(limits = c(-0.05, NA), name = "Mantel correlation") +
  scale_x_continuous(breaks = c(0, 3, 6, 9, 12)) +
  scale_color_discrete(name = "Time lag") +
  xlab("Distance (m)") +
  ggnomics::facet_nested(metric + tech + amplicon ~ guild, scales = "fixed", nest_line = TRUE) +
  scale_shape_manual(values = c("TRUE" = 16, "FALSE" = 1)) +
  theme(strip.background = element_blank())
```

@@

```{r spatio-temporal, fig.cap = "(ref:spatio)"}
variog_data <- light_plan %>%
  filter(step == "variofitST") %>%
  select(variofitST, variogST, guild, metric, tech, amplicon) %>%
  mutate_at(c("guild", "metric", "tech", "amplicon"), str_replace_all, "\"", "") %>%
  mutate(
    guild = plyr::mapvalues(guild, c("fungi", "ecm"), c("All Fungi", "ECM")),
    metric = plyr::mapvalues(metric, c("bray", "wunifrac"), c("Bray-Curtis", "W-UNIFRAC")),
    amplicon = factor(
      amplicon,
      levels = c("Short", "Long")
    )
  ) %>%
  mutate_at(
    c("variogST", "variofitST"),
    lapply,
    readd,
    character_only = TRUE,
    cache = cache)

@@

variog_empir %>%
  ggplot(aes(dist, gamma, color = timelag, size = n)) +
  # geom_hline(data = variog_sills, aes(yintercept = sills), linetype = "dashed", color = "gray50") +
  # geom_vline(data = variog_data, aes(xintercept = range), linetype = "dashed", color = "gray50") +
  # geom_jitter(shape = 46, alpha = 0.5) +
  # geom_line(size = 1.5, alpha = 0.5) +
  geom_point(alpha = 0.5) +
  # geom_smooth(data = variog_points, aes(fill = timelag), color = FALSE, alpha = 0.5) +
  geom_line(data = filter(variog_fit, timelag == 1), size = 0.5, color = "darkcyan", alpha = 0.8) +
  geom_line(data = filter(variog_fit, timelag == 0), size = 0.5, color = "red1", alpha = 0.8) +
  scale_x_continuous(expand = expand_scale(0, 0), limits = c(0, 25), name = "Distance (m)") +
  scale_color_discrete(name = "Time lag") +
  scale_size_continuous(name = "Comparisons") +
  scale_y_continuous(name = "Community dissimilarity") +
  ggnomics::facet_nested(metric + tech + amplicon ~ guild, scales = "free", nest_line = TRUE) +
  theme(strip.background = element_blank())

@@

(ref:full-length) Comparison of the length of sequences from different length amplicons and sequencing technologies.  Length distribution (*a*, *c*) and empirical cumulative distribution function (*b*, *d*) for short (*a*, *b*) and long (*c*, *d*) amplicons, respectively.

```{r full-length, fig.cap = "(ref:full-length)", fig.height = 6}
full_dens <- readd(reads_table, cache = cache) %>%
  dplyr::left_join(dplyr::select(datasets, seq_run, tech, amplicon)) %>%
  dplyr::filter(region %in% c("long", "short")) %>%
  dplyr::group_by(seq_run) %>%
  dplyr::mutate_at("amplicon", factor, levels = c("Short", "Long")) %>%
  dplyr::mutate_at("tech", factor, levels = c("PacBio", "Illumina", "Ion Torrent")) %>%
  dplyr::mutate(
    reads = reads / sum(reads)
  )

full_dens_short <- full_dens %>%
  filter(amplicon == "Short") %>%
  ggplot(aes(length, weight = reads, color = tech, group = tech)) +
  stat_density(bw = 0.5, geom = "line", position = "identity") +
  scale_color_brewer(type = "qual", guide = "none", palette = "Set2") +
  scale_y_continuous(name = "Density") +
  scale_x_continuous(name = NULL, labels = NULL, breaks = NULL) +
  ggnomics::facet_nested(rows = vars(tech), scales = "free_y") +
  theme(strip.background = element_blank())

full_dens_long <- full_dens %>%
  filter(amplicon == "Long") %>%
  ggplot(aes(length, weight = reads, color = tech, group = tech)) +
  stat_density(bw = 0.5, geom = "line", position = "identity") +
  scale_color_brewer(type = "qual", guide = "none", palette = "Set2") +
  scale_y_continuous(name = "Density") +
  scale_x_continuous(name = NULL, labels = NULL, breaks = NULL) +
  ggnomics::facet_nested(rows = vars(tech), scales = "free_y") +
  theme(strip.background = element_blank())

full_ecdf <- readd(reads_table, cache = cache) %>%
  dplyr::filter(region  %in% c("long", "short")) %>%
  left_join(select(datasets, seq_run, amplicon, tech, machine), by = "seq_run") %>%
  mutate_at("tech", factor, levels = c("PacBio", "Illumina", "Ion Torrent")) %>%
  dplyr::group_by(seq_run) %>%
  dplyr::mutate(reads = reads / sum(reads))
